Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-22 cav2017/cav2017.tex
text

Asymptotic stability boils down to Lyapunov stability and convergence to an equilibrium point \todo{add reference}.  
Lyapunov stability requires that model trajectories remain in a neighbourhood of the equilibrium point. 
In the case of closed-loop LTI models as in \eqref{eq:closedloopss}, 
the equilibrium point is the origin (if the reference is equal to zero), or an offset thereof.

@@

The states values must be computed considering the digital controller error, i.e., the error introduced by the ADC and
digital controller, and the digital reference signal. In this paper, we consider systems with
a reference signal equal to zero. 
%the FWL effects in all operations related to the control action~\eqref{eq:controlaction} 
%and disregarding it in the state equation~\eqref{eq:plant}.
Thus, the states are computed using the following equations. We use $\mathcal{F}_{\langle I_c,F_c \rangle}$ to 
to denote numbers represented with the precision of the digital controller. All other numbers
are represented with the precision of our plant model:
%
\begin{align*}
u_{k}&=-(\mathcal{F}_{\langle I_c,F_c \rangle}(K)\cdot\mathcal{F}_{\langle I_c,F_c \rangle}(x_{k})) \\
x_{k+1}&=Ax_{k}+Bu_{k}
\end{align*}
%\left\lbrace
%\begin{array}{l}
%u_{k}=\mathcal{F}_{\langle I,F \rangle}(r_{k}-\mathcal{F}_{\langle I,F \rangle}(\mathcal{F}_{\langle I,F \rangle}(K)\cdot\mathcal{F}_{\langle I,F \rangle}(x_{k})))\\
%x_{k+1}=Ax_{k}+Bu_{k}
%\end{array}\right.
%\end{equation}

@@

The systems and models we consider have two sources of error due to numeric representation; the numeric error introduced by 
the fixed point numbers used to model our plant, i.e, the plant dynamics $A_d$, $B_d$ and the states; 
and the numeric error introduced by the digital controller, which uses an analogue-to-digital converter to read
the states as a fixed-point number and performs operations in this domain. In this section we formally outline the errors introduced by
fixed point representation of numbers; in further sections we will differentiate between these
two errors by referring to them as the \emph{plant error} and the \emph{digital controller error}.

Let $\mathbb{R} \langle I,F \rangle$ be a fixed point domain with $I$ bits representing the integer part and $F$ bits
representing the decimal part. The smallest number that can be represented in this domain is $c_m=2^{-F}$.
The following approximation errors will arise:
\begin{enumerate}
\item {\bf Truncation:}
Let $\mathcal{F}\langle I,F \rangle(\cdot) : \mathbb{R} \rightarrow \mathbb{R}\langle I,F \rangle, \tilde x=\mathcal{F}\langle I,F \rangle(x) = x-\delta_T : \delta_T=x\ \%_{c_m}\ \tilde x$, where $\%_{c_m}$ is the modulus operation performed wrt the last bit.
$\delta_T$ is the truncation error and it will propagate across operations.
\item {\bf Rounding:}
The following errors appear in simple operations:
\begin{enumerate}
\item Addition/Substraction: these operations only propagate errors coming from truncation of the operands.
$c_1\pm c_2= c_3 + \delta_3 :|\delta_3| \leq |\delta_{T1}| + |\delta_{T2}|$ 
\item Multiplication: $c_1 * c_2= c_3 + \delta_3 :|\delta_3| \leq |\delta_{T1}*c_2| + |\delta_{T2}*c_1| + c_m$ 
\item Division will not be applied in these implementations.
\end{enumerate}

>>>>

Asymptotic stability boils down to Lyapunov stability and convergence to an equilibrium point \todo{add reference}.  
Lyapunov stability requires that model trajectories remain in a neighbourhood of the equilibrium point. 
In the case of closed-loop LTI models as in \eqref{eq:closedloopss}, 
the equilibrium point is the origin (if the reference is equal to zero), or an offset thereof.  
In this paper, we consider systems with a reference signal equal to zero.

@@

@@

\todo{I've moved the following paragraph to this section, from that on safety specs.}
The states values must be computed considering the digital controller error, i.e., the error introduced by the ADC and the digital controller. 
%, and the digital reference signal. 
%In this paper, we consider systems with a reference signal equal to zero. 
%the FWL effects in all operations related to the control action~\eqref{eq:controlaction} 
%and disregarding it in the state equation~\eqref{eq:plant}.
Thus, the states are computed using the following equations. We use $\mathcal{F}_{\langle I_c,F_c \rangle}$ \todo{explain and clarify notation.}
to to denote numbers represented with the precision of the digital controller. All other numbers
are represented with the precision of our plant model:
%
\begin{align*}
u_{k}&=-(\mathcal{F}_{\langle I_c,F_c \rangle}(K)\cdot\mathcal{F}_{\langle I_c,F_c \rangle}(x_{k})).  
%\\ x_{k+1}&=A_d x_{k} + B_du_{k}
\end{align*}
%\left\lbrace
%\begin{array}{l}
%u_{k}=\mathcal{F}_{\langle I,F \rangle}(r_{k}-\mathcal{F}_{\langle I,F \rangle}(\mathcal{F}_{\langle I,F \rangle}(K)\cdot\mathcal{F}_{\langle I,F \rangle}(x_{k})))\\
%x_{k+1}=Ax_{k}+Bu_{k}
%\end{array}\right.
%\end{equation}

\bigskip

The 
%systems and 
models we consider have two sources of error due to numeric representation. 
The first is the numeric error introduced by the fixed-point numbers employed to model the plant, i.e. to represent the plant dynamics $A_d$, $B_d$ and its states. 
The second is the numeric error introduced by the digital controller, 
which 
%uses an analogue-to-digital converter to read the states as a fixed-point number and performs operations in this domain. 
performs operations in the domain of fixed-point numbers. 
In this section we formally outline the errors introduced by fixed-point representation of numbers. 
In further sections we will differentiate between these two errors, 
by referring to them as the \emph{plant error} and the \emph{digital controller error} respectively.

Let $\mathbb{R} \langle I,F \rangle$ \todo{notice this notation differs from the one used above} be a fixed point domain with $I$ bits representing the integer part and $F$ bits
representing the decimal part. The smallest number that can be represented in this domain is $c_m=2^{-F}$.
The following approximation errors will arise:
\begin{enumerate}
\item {\bf Truncation:}
Let $\mathcal{F}\langle I,F \rangle(\cdot) : \mathbb{R} \rightarrow \mathbb{R}\langle I,F \rangle, \tilde x=\mathcal{F}\langle I,F \rangle(x) = x-\delta_T : \delta_T=x\ \%_{c_m}\ \tilde x$, where $\%_{c_m}$ is the modulus operation performed wrt the last bit.
$\delta_T$ is the truncation error and it will propagate across operations.
\item {\bf Rounding:}
The following errors appear in simple operations:
\begin{enumerate}
\item Addition/Substraction: these operations only propagate errors coming from truncation of the operands.
$c_1\pm c_2= c_3 + \delta_3 :|\delta_3| \leq |\delta_{T1}| + |\delta_{T2}|$ 
\item Multiplication: $c_1 * c_2= c_3 + \delta_3 :|\delta_3| \leq |\delta_{T1}*c_2| + |\delta_{T2}*c_1| + c_m$ 
\item Division will not be applied in these implementations.
\end{enumerate}
